Viscous Transport in Eroding Porous Media

### 2019-09-30 docs/transport/erosion_transport.tex
NM

To compute stable simulations of erosion, we use high-order methods in
both space and time. The time integration method is unchanged from our
previous work~\cite{qua-moo2018}.  We apply a mild regularization and a
smoothing term to eliminate numerical instabilities caused by changes in
sign of the shear stress, and we use a stable second-order Runge-Kutta
method applied to the {\thL} coordinates~\cite{hou-low-she1994} of the
eroding grains.  In this work, we do introduce a new quadrature method
to resolve dense suspensions of eroding grains.  The accuracy of the
trapezoid rule, which we used in our previous work, is sufficient for
bodies that are sufficiently separated~\cite{tre-wei2014}. However, it
is inaccurate when two grains are nearly-touching.  One of the earliest
works that addresses quadrature methods for nearly-singular integrands
was developed by Baker and Shelley~\cite{bak-she1986}, and in recent
years, many other near-singular integration schemes have been
developed~\cite{oja-tor2015, kli-tor2018, hel-oja2008a, bea-yin-wil2016,
bea-lai2001, klo-bar-gre-one2013}.  We use a Barycentric quadrature
rule~\cite{bar2014, bar-wu-vee2015} since it is a non-intrusive
modification of the trapezoid rule, and the error is guaranteed to be
uniformly bounded.  We extend the original quadrature method so that it
can be used to compute the gradient of the velocity that is required to
compute the shear stress and the fluid vorticity.

@@

To characterize the transport, particle trajectories need to be
computed. Depending on the application, microscale transport can be
modelled as pure advection~\cite{dea-qua-bir-jua2018,
leb-ded-dav-bou2007, cve-che-wen1996, puy-gou-den2019},
advection-diffusion~\cite{cus-hu-den1995, dag1987, den-ica-hid2018}, or
with a random walk~\cite{saf1959, bij-blu2006, ber-sch-sil2000}.  In
this work, we assume that the trajectories are governed purely by
advection (ie.~no diffusion). Therefore, the particle trajectories are
identical to the streamlines, and they can be found, for example, with a
continuous time random walk~\cite{mon-wei1965, den-cor-sch-ber2004,
leb-den-car2008, ber-cor-den-sch2006}.  Instead,  we compute
trajectories $\ss(t)$ that are initialized at $\ss_0$ by solving
\begin{align}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0,
  \label{eqn:tracers}
\end{align}
where $\uu$ is the fluid velocity.  Since there is no stiff diffusive
term in~\eqref{eqn:tracers}, we use a fourth-order explicit Runge-Kutta
time stepping method.  In addition, the Barycentric quadrature rule is
used to accurately compute trajectories that are close to an eroding
grain.

>>>>

To compute stable simulations of erosion, we use high-order methods in
both space and time. The time integration method is unchanged from our
previous work~\cite{qua-moo2018}.  We apply a mild regularization and a
smoothing term to eliminate numerical instabilities caused by changes in
sign of the shear stress, and we use a stable second-order Runge-Kutta
method applied to the {\thL} coordinates~\cite{hou-low-she1994} of the
eroding grains.  In this work, we \st{do} introduce a new quadrature method
to resolve dense suspensions of eroding grains.  The accuracy of the
trapezoid rule, which we used in our previous work, is \st{sufficient} \nick{adequate} for bodies that are sufficiently separated~\cite{tre-wei2014}. 
\todo[inline]{I changed the above to `adequate' so that `sufficient' would not be used twice in a row.} 
However, it \nick{loses accuracy} whenever \st{two} grains are nearly-touching.  One of the earliest works \nick{to address} quadrature methods for nearly-singular integrands was developed by Baker and Shelley~\cite{bak-she1986}, and in recent years, many other near-singular integration schemes have been \nick{followed}~\cite{oja-tor2015, kli-tor2018, hel-oja2008a, bea-yin-wil2016, bea-lai2001, klo-bar-gre-one2013}.  We use a Barycentric quadrature
\nick{method}~\cite{bar2014, bar-wu-vee2015} since it is a non-intrusive
modification of the trapezoid rule, and the error is guaranteed to be
uniformly bounded.  We extend the original quadrature method \nick{to compute the velocity gradient, which is needed to evaluate shear stress and fluid vorticity}.

@@

To characterize the transport, particle trajectories \nick{must} be
computed. Depending on the application, microscale transport can be
modelled as pure advection~\cite{dea-qua-bir-jua2018,
leb-ded-dav-bou2007, cve-che-wen1996, puy-gou-den2019},
advection-diffusion~\cite{cus-hu-den1995, dag1987, den-ica-hid2018}, or
with a random walk~\cite{saf1959, bij-blu2006, ber-sch-sil2000}.  In
this work, we assume trajectories \nick{to be governed by pure advection} (ie.~no diffusion). Therefore, the particle trajectories are
identical to the streamlines, and they can be found, for example, with a
continuous time random walk~\cite{mon-wei1965, den-cor-sch-ber2004,
leb-den-car2008, ber-cor-den-sch2006}.  
\todo[inline]{The logic from the sentence above to the one below seems awkward to me. The above sentence makes me think that a continuous time random-walk will be used in our paper. Then, suddenly, we say we are going to do something else instead. Maybe we can kill the above sentence and dump the references (if you want to keep them) into the earlier sentence that mentions a random walk? Or maybe just a slight rephrasing would make the transition less awkward.}
Instead,  we compute
trajectories $\ss(t)$ that are initialized at $\ss_0$ by solving
\begin{align}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0,
  \label{eqn:tracers}
\end{align}
where $\uu$ is the fluid velocity.  Since there is no stiff diffusive
term in~\eqref{eqn:tracers}, we use a fourth-order explicit Runge-Kutta
time stepping method.  In addition, the Barycentric quadrature rule is
used to accurately compute trajectories that are close to an eroding
grain.
